\documentclass{tropical}
\title{Plank Countdown}
\author{Nyan}
\date\today
\begin{document}
\begin{frame}
\maketitle
\end{frame}
\begin{frame}{Sample problem}
Please do not actually try this one manually.
\begin{block}{Problem 0}
What are the first four digits of $3^{15}$?
\end{block}
\end{frame}

\begin{frame}{Answer 0}
Source: well known joke in the math community.
Answer: 1434
\end{frame}

\begin{frame}{Problem 1}
Let $a$ and $b$ be five-digit palindromes 
(without leading zeroes) such that $a<b$
and there are no other
five-digit palindromes strictly between $a$ and $b$.
What are all possible values of $b-a$? 
(A number is a palindrome if it reads 
the same forwards and backwards in base 10.)
\end{frame}

\begin{frame}{Answer 1}
Casework on how many `carries' there are between consecutives.
$100$, $110$, $11$.\\
Source: HMMT Feb 2018 C2.
\end{frame}

\begin{frame}{Problem 2}
Quadrilateral $ABCD$ is cyclic with $AB = CD = 6$. Given that $AC = BD = 8$ and $AD+3 = BC$, what is $[ABCD]?$\\
\begin{block}{Note}
$[A_1\dots A_n]$ denotes the area of that polygon.
\end{block}
\end{frame}
\begin{frame}{Answer 2}
Ptolemy to find the missing lengths $AD$, $BC$, then Brahmagupta to find the area. $33\sqrt{15}/4$.
\\
Source: BMT 2020 TB G2.
\end{frame}
\begin{frame}{Problem 3}
Every face of a cube is colored one of $3$ colors at random. What is the expected number of edges
that lie along two faces of different colors?
\end{frame}
\begin{frame}{Answer 3}
Linearity of expectation. Probability that two adjacent faces are different colours is $2/3$. $8$.
\\
Source: BMT 2018 TB C1.
\begin{block}{Remark}
BMT tiebreaker problems are also meant to be fast paced- 
these problems are supposed to be solved in 
$15$ mins though most contestants finish before then. 
CMM TB does not work for this purpose, on the other hand....
\end{block}
\end{frame}
\begin{frame}{Problem 4}
There is a unique 4-digit positive integer
$n=\underline{abcd}$ with $\underline{ab}\cdot\underline{cd} =n/3-2$.
What is the sum of the digits
of $n$?
\end{frame}
\begin{frame}{Answer 4}
$12$. This is a 1434 joke invented by one of my friends.
Source: 2023 AMC 12B*/13 (a mock AMC you can find on AoPS)
\end{frame}
\begin{frame}{Problem 5}
Find the number of ordered pairs of integers $(a,b)$ such that $a,b$ are divisors of 720 
but $ab$ is not.
\end{frame}
\begin{frame}{Answer 5}
Complementary counting. 2520.\\
Source: HMMT Feb 2016 C3
\end{frame}
\begin{frame}{Problem 6}
Compute the probability that a random permutation of the letters in BERKELEY does not
have the three E’s all on the same side of the Y.
\end{frame}
\begin{frame}{Answer 6}
Shockingly this is a BMT problem again. (2018 TB C1)
We only care about the locations of the Y's and the E's. $2\cdot 1/4=1/2$.
\end{frame}
\begin{frame}{Problem 7}
Right triangle $ABC$ with its right angle at $B$ has angle bisector $\ol{AD}$ 
with $D$ on $\ol{BC}$, as well
as altitude $\ol{BE}$ with $E$ on $\ol{AC}$. 
If $\ol{DE} \perp\ol{BC}$ and $AB = 1$, compute $AC$.
\end{frame}
\begin{frame}{Answer 7}
Trig bash. $(1+\sqrt5)/2$.\\
Source: BMT 2021 TB G2
\end{frame}
\begin{frame}{Problem 8}
Let $\Omega$ and $\omega$ be circles with radii 123 and 61, respectively, 
such that the center of $\Omega$ lies on $\omega$. 
A chord of $\Omega$ is cut by $\omega$ into three segments, 
whose lengths are in the ratio $1 : 2 : 3$ in that order. 
Given that
this chord is not a diameter of $\omega$, compute the length of this chord.
\end{frame}
\begin{frame}{Answer 8}
Synthetic and power of a point. $42$.\\
Source: 2024 HMMT Feb G3
\end{frame}
\begin{frame}{Problem 9}
In triangle $ABC$ with $AB = 8$ and $AC = 10$, 
the incenter $I$ is reflected across side $AB$ to point $X$ 
and across side $AC$ to point $Y$. 
Given that segment $XY$ bisects $\ol{AI}$, compute $BC^2$. 
\end{frame}
\begin{frame}{Answer 9}
Use synthetic to get $\angle A=60^\circ$. $84$ by some bashing.\\
Source: HMMT Nov 2020 General 7
\end{frame}
\begin{frame}{Problem 10}
Find the integer closest to 
\[\frac1{\sqrt[4]{5^4+1}-\sqrt[4]{5^4-1}}.\]
\end{frame}
\begin{frame}{Answer 10}
Rationalise the denominator. 250.\\
Source: HMMT Feb 2014 A6
\end{frame}
\begin{frame}{Problem 11}
Greta writes $2, 3, . . . , 101$ on a chalkboard. 
Every minute she erases two chosen numbers $x,y$ from the board,
and writes $xy/(x+y-1)$ in their place. 
After 99 minutes, what number remains?
\end{frame}
\begin{frame}{Answer 11}
This is a contrived invariant problem. The invariant in question is 
\[\prod_k \frac1{1-1/x_k}\]
where the nums on the board are $x_1,...$.\\
Source: BMT 2022 TB A3
\end{frame}
\begin{frame}{Problem 12}
A positive integer is called extra-even if all of its digits are even. Compute the number of positive
integers n less than or equal to 2022 such that both n and 2n are both extra-even.
\end{frame}
\begin{frame}{Answer 12}
Figure out from experimenting that $0,2,4$ are the only legal digits. $31$.\\
Source: BMT 2022 TB C2
\end{frame}
\begin{frame}{Problem}
Let a and b be real numbers such that $a + b = 16$ and $a^3 - b^3 = ab^2 - a^2b + 2048$. 
Compute $ab^2$.
\end{frame}
\begin{frame}{Answer}
Factor, divide, etc to obtain $a,b$. $192$.\\
Source: BMT 2025 TB A2
\end{frame}
\begin{frame}{I ran out of problem ideas}
Thank you for playing!! :3
\end{frame}
\end{document}
